---
title: "Reproducible Research: Peer Assessment 1, by Shuai Wang"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
raw_data<-read.csv('activity.csv')
summary(raw_data)
```
## What is mean total number of steps taken per day?
1. ignore the missing data using na.omit
2. use reshape2 library to sum by date
3. plot the histogram
```{r}
complete_data<-na.omit(raw_data)
complete_data$date<-as.Date(complete_data$date)
complete_data$interval<-as.factor(complete_data$interval)
aggdata <-aggregate(steps~date+interval,complete_data,FUN=mean)
print(aggdata)
```
Plot the histogram
```{r}
hist(aggdata$steps, main="Histogram of steps taken per day",xlab="Number of Steps", ylab="Count")
```
So we can easily report mean and median by  
```{r}
print(mean(aggdata$steps))
print(median(aggdata$steps))
```
## What is the average daily activity pattern?
1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

``` 

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
